---
title: "3.2 Stability analysis (glmnet) with normalized data"
author: "Nicolas Bennett"
date: "`r Sys.Date()`"
output:
  rmarkdown::pdf_document:
    includes:
      in_header: 0_header.tex 
vignette: >
  %\VignetteIndexEntry{3.2 Stability analysis (glmnet) with normalized data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, setup, echo=FALSE}
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(singleCellFeatures))
suppressPackageStartupMessages(library(xtable))
options(scipen=1, digits=2, singleCellFeatures.progressBars="none")
opts_chunk$set(cache.extra = rand_seed)
```

```{r, loadData-mtor, cache=TRUE, collapse=TRUE}
suppressPackageStartupMessages(library(singleCellFeatures))
suppressPackageStartupMessages(library(xtable))
mtor <- findWells(contents="MTOR", experiment="brucella-du-k[12]")
scra <- findWells(well.names="H2", plates=sapply(mtor, getBarcode))

data.mtor <- unlist(getSingleCellData(c(mtor, scra)), recursive=FALSE)
data.mtor <- lapply(data.mtor, cleanData, "lower")
data.mtor <- makeFeatureCompatible(data.mtor)
reps  <- 100

```

To find out if good predition/data separation is only given in special situations or occurs readily, stability analysis using glmnet is performed on the *MTOR* (H6)/*SCRAMBLED* (H2) pair and on several randomly selected well pairs on the same plates.

First, wells containing siRNA for the gene *MTOR* are searched for within the kinome-wide Dharmacon unpooled screens (replicates 1 and 2). Then on the plates containing those wells, scrambled control experiments are looked up (in well H2). The data for the resulting 16 wells is loaded, cleaned up (lower 5% quantile in terms of well-level cell counts is discarded) and melted into data frames.

```{r, stability-mtor, cache=TRUE, collapse=TRUE, warning=FALSE}
mtor.si <- lapply(lapply(mtor, getBarcode), function(bc, data.mtor, n) {
  glmBootstrapStability(data.mtor[[paste0(bc, ".H2")]],
                        data.mtor[[paste0(bc, ".H6")]],
                        n.rep=n, norm.feat="intensity",
                        norm.method="unitInterval", alpha=0.5)
}, data.mtor, reps)

mtor.k1 <- glmBootstrapStability(
  data.mtor[grep("-2C.H2$", names(data.mtor))],
  data.mtor[grep("-2C.H6$", names(data.mtor))],
  n.rep=reps, norm.feat="intensity", norm.method="unitInterval", alpha=0.5)
mtor.k2 <- glmBootstrapStability(
  data.mtor[grep("-2D.H2$", names(data.mtor))],
  data.mtor[grep("-2D.H6$", names(data.mtor))],
  n.rep=reps, norm.feat="intensity", norm.method="unitInterval", alpha=0.5)
mtor.al <- glmBootstrapStability(data.mtor[grep(".H2$", names(data.mtor))],
                                 data.mtor[grep(".H6$", names(data.mtor))],
                                 n.rep=reps, norm.feat="intensity",
                                 norm.method="unitInterval", alpha=0.5)

```

\newpage
\blandscape

```{r, table-mtor, results="asis", warning=FALSE, echo=FALSE}
table1 <- Reduce(function(x, y) {
  res <- merge(x, y, all=TRUE, by=0)
  rownames(res) <- res$Row.names
  res$Row.names <- NULL
  return(res)
}, c(mtor.si, list(mtor.k1, mtor.k2, mtor.al)), accumulate=FALSE)

colnames(table1) <- c(sapply(mtor, getBarcode), "K1", "K2",
                            "DU")
table1[is.na(table1)] <- 0

table1$sums <- rowSums(table1)
table1 <- table1[order(table1$sums, decreasing=TRUE),]
print(xtable(head(table1, n=50), digits=0), size="footnotesize", comment=FALSE)

```

\elandscape
\newpage

Stability analysis on *MTOR* (Well H6) versus *SCRAMBLED* (Well H2) is performed with 100 resampling iterations, each run on 70% of the data. The top 20 coefficients of each iteration are selected and their frequencies are tabulated. K1 consists of merged \*-2C wells, K2 corresponds to \*-2D and DU encompasses all 8 well pairs. Models were fit with glmnet, and default parameters (elastic net penalty: $$\alpha = 0.5$$).

```{r, loadData-rand, cache=TRUE, collapse=TRUE}
barcodes <- sapply(findWells(contents="MTOR", experiment="brucella-du-k[12]"),
                   getBarcode)
set.seed(9)
wells <- sapply(barcodes, function(bc) {
  loc <- paste0(sample(LETTERS[1:16], 2), sample(1:24, 2))
  findWells(plates=bc, well.names=loc)
})

data.rand  <- unlist(getSingleCellData(wells), recursive=FALSE)
data.rand  <- lapply(data.rand, cleanData, "lower")
data.rand  <- makeFeatureCompatible(data.rand)

```

For each plate holding an *MTOR* well, a randomly selected pair of wells is loaded and prepared for analysis. This yields comparisons between control well pairs, siRNA well pairs and control/siRNA well pairs. 

```{r, stability-rand, cache=TRUE, warning=FALSE, collapse=TRUE}
rand.si <- lapply(barcodes, function(bc, data, n) {
  plate <- grep(bc, names(data))
  glmBootstrapStability(data[[plate[1]]], data[[plate[2]]],
                        n.rep=n, norm.feat="intensity",
                        norm.method="unitInterval", alpha=0.5)
}, data.rand, reps)

```

For each plate holding an *MTOR* well, a randomly selected pari of wells is loaded and prepared for analysis. This yields comparisons between control well pairs, siRNA well pairs and control/siRNA well pairs. 

\newpage
\blandscape

```{r, table-rand, results="asis", warning=FALSE, echo=FALSE}
table2 <- Reduce(function(x, y) {
  res <- merge(x, y, all=TRUE, by=0)
  rownames(res) <- res$Row.names
  res$Row.names <- NULL
  return(res)
}, rand.si, accumulate=FALSE)

colnames(table2) <- barcodes
table2[is.na(table2)] <- 0

table2$sums <- rowSums(table2)
table2 <- table2[order(table2$sums, decreasing=TRUE),]
print(xtable(head(table2, n=50), digits=0), size="footnotesize", comment=FALSE)

```

\elandscape
\newpage

The randomply selected well pairs are `r paste0(unlist(lapply(wells, getWellName))[seq(from=1, by=2, length.out=8)], "/", unlist(lapply(wells, getWellName))[seq(from=2, by=2, length.out=8)], " (", barcodes, ")", collapse=", ")`. Bootstrapping is done with 100 resampling iterations, each run on 70% of the data. The top 20 coefficients of each iteration are selected and their frequencies are tabulated. Models are fit with glmnet, and default parameters (elastic net penalty: $\alpha = 0.5$).

The randomply generated table compares well to the *MTOR* table and the dominance of intensity features is striking. This suggests, some kind of normalisazion to be required for sensible model fits.
